---
title: "exp1_analysis"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(ggplot2)
library(jsonlite)
library(tidyr)
library(stringr)
library(dplyr)
library(ggthemes)
library(ggrepel)
library(dichromat)
library(lme4)
library(lmerTest)
library(performance)
library(languageR)
library(gghighlight)
theme_set(theme_bw())
library(brms)
library(rstan)
# library(plot3D)


setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
# color-blind-friendly palette
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7") 
```

```{r import data}
top_words_selected <- read.csv("../explore/top_words_selected.csv")
top_words_selected <- select(top_words_selected, "Word", "voice", "valence_group", "valence_bin", "valence_bin_group", "arousal_bin", "valence_arousal_group", "carousal_mean", "crelativeValence", "valence_mean", "mean_residuals_abs","mean_acceptability_norm")
data_raw <- read.csv("../../Submiterator/exp1/exp1-trials.csv")
data_participants_raw <- read.csv("../../Submiterator/exp1/exp1-subject_information.csv")
data_at_issue <- read.csv("../../data/atissue_data-bypred.csv")
frequencies = read_csv("../../data/frequencies.csv")
anand = read_csv("../../data/anand-etal-typology_filledout.csv")
names(anand)[names(anand) == "predicate"] <- "Word"
names(anand)[names(anand) == "broad class"] <- "broad_class"
bert_score = read_csv("../../data/bert_scores.csv")
```

# Native speakers
## We exclude any speakers who did not report English as a native language in the post-survey demographic questions.
```{r filter native English speakers, warning = FALSE}
# get participants whose native language is English
data_participants <- data_participants_raw %>%
  mutate(language = tolower(language))

#filter out participants who did not rep ort English as native language
data_participants = filter(data_participants, grepl("eng", language))

worker_id <- data_participants$workerid

data_select <- data_raw %>%
  filter(utterance != "bot_check") %>%
  filter(workerid %in% worker_id)

data_select <- data_select %>%
  mutate(response_num = as.numeric(response))
  #mutate(response_num = as.numeric(levels(response))[response])
```
# Control trials
## We exclude any speakers whose mean performance on control trials is 3 SD above the group mean.
```{r filter out control trials}
data_control <- data_select %>%
  filter(exp == "control") 

ggplot(data_control, aes(x=response_num)) +
  geom_histogram() +
  facet_wrap(~utterance)

worker_control = ggplot(data_control, aes(x=response_num)) +
  geom_histogram() +
  facet_wrap(~workerid)
# ggsave(worker_control, file="graphs/worker_control_items.pdf",width=15,height=20)

control_mean = mean(data_control[, "response_num"])
control_median = median(data_control[, "response_num"])
control_sd = sd(data_control[, "response_num"])
max_control = control_mean + 3 * control_sd

# omit mean response 3 SD above the group mean (per factive paper)
id_omit <- data_control %>%
  group_by(workerid) %>%
  summarise(response_mean_control = mean(response_num)) %>%
  filter(max_control < response_mean_control)
id_omit <- id_omit[, "workerid"]

id_select <- data_control %>%
  group_by(workerid) %>%
  summarise(response_mean_control = mean(response_num)) %>%
  filter(response_mean_control <= max_control)

id_select <- pull(id_select[c("workerid")])
```

```{r filter out stimuli trials}
data_stim <- data_select %>% 
  filter(workerid %in% id_select) %>%
  filter(exp == "stim") %>%
  mutate(Word = ifelse(Word == "facinate", "fascinate", as.character(Word))) ## coded "fascinate" wrongly :(

data <- merge(data_stim, top_words_selected, by = "Word")

predlevels = unique(data$Word) %>% 
  sort()
# JD: hm, "confess" and "remark" are in frequencies but not in data, wonder why?
frequencies = frequencies %>% 
  filter(! Word %in% c("confess","remark")) %>% 
  arrange(Word)
frequencies$Word = predlevels # risky new assignment if predicates not ordered correctly
data = data %>% 
  left_join(data_at_issue,by=c("Word")) %>% 
  left_join(frequencies, by=c("Word"))

# Add in the Anand et al (2019) coding
data = left_join(data,anand, by=c("Word"))

bert_score$BertSurprisalCC = -log(bert_score$BertScoreCC)
bert_score$BertSurprisalNoCC = -log(bert_score$BertScoreNoCC)
bert_score$cBertSurprisalCC = bert_score$BertSurprisalCC - mean(bert_score$BertSurprisalCC)
bert_score$cBertSurprisalNoCC = bert_score$BertSurprisalNoCC - mean(bert_score$BertSurprisalNoCC)

avg_bert = bert_score %>%
  group_by(Verb) %>%
  summarize(cmeanBertSurprisalCC = mean(cBertSurprisalCC),cmeanBertSurprisalNoCC = mean(cBertSurprisalNoCC))
names(avg_bert)[names(avg_bert) == "Verb"] <- "Word"
data = left_join(data,avg_bert, by=c("Word"))


```


# Mixed effects linear model and ANOVA
```{r mixed effects linear model, include = FALSE, warning = FALSE, echo=FALSE}
m = lmer(response_num ~ carousal_mean * crelativeValence + (1 + (carousal_mean * crelativeValence) | workerid) + (1 | Word) + (1 + (carousal_mean * crelativeValence) | content), data = data)
summary(m) # main effects of arousal and relative valence!
plot(fitted(m), residuals(m))

#pdf(file="modelcheck_linear.pdf",height=8,width=9)
#check_model(m)
#dev.off()
```

```{r anova, include = FALSE, warning = FALSE, echo=FALSE}
projectivity.full = lmer(response_num ~ carousal_mean * crelativeValence + (1 + (carousal_mean * crelativeValence) | workerid) + (1 | Word) + (1 + (carousal_mean * crelativeValence) | content), data = data_model, REML = FALSE)
projectivity.reduced = lmer(response_num ~ 1 + (1 + (carousal_mean * crelativeValence) | workerid) + (1 | Word) + (1 + (carousal_mean * crelativeValence) | content), data = data, REML = FALSE)
anova(projectivity.reduced, projectivity.full)


#interaction removed
projectivity.int = lmer(response_num ~ carousal_mean + crelativeValence + (1 + (carousal_mean * crelativeValence) | workerid) + (1 | Word) + (1 + (carousal_mean * crelativeValence) | content), data = data, REML = FALSE)
anova(projectivity.int, projectivity.full)

#only arousal
projectivity.arousal = lmer(response_num ~ carousal_mean + (1 + (carousal_mean * crelativeValence) | workerid) + (1 | Word) + (1 + (carousal_mean * crelativeValence) | content), data = data, REML = FALSE)
anova(projectivity.int, projectivity.arousal)

#only valence
projectivity.valence = lmer(response_num ~ crelativeValence + (1 + (carousal_mean * crelativeValence) | workerid) + (1 | Word) + (1 + (carousal_mean * crelativeValence) | content), data = data, REML = FALSE)
anova(projectivity.int, projectivity.valence)

```

```{r mixed bayes, include = FALSE, echo = FALSE}

# m.bayes = brm(response_num ~ carousal_mean * crelativeValence + (1 + (carousal_mean * crelativeValence) | workerid) + (1 | Word) + (1 + (carousal_mean * crelativeValence) | content), 
#               data = data_model,
#               chains = 4,
#               cores = 4)
# saveRDS(m.bayes, "bayes.rds")
m.bayes <- readRDS("bayes.rds")
summary(m.bayes) # 0 is in 95% CI for interaction, but not for valence / arousal separately
```


# Main analysis reported by JD in workshop talk

```{r}
data$workerid = as.factor(as.character(data$workerid))
data$catissue = data$mean_atissue - mean(data$mean_atissue)
# Surprisal from google n-gram
data$cSurprisal = data$Surprisal - mean(data$Surprisal)

```

# Surprisal
## Google Books NGram 
```{r mixed effects linear model, include = FALSE, warning = FALSE, echo=FALSE}
m.full <- lmer(response_num ~ carousal_mean*crelativeValence + catissue + cSurprisal + (1 | Word) + (1 | content) + (1|workerid), data = data, REML = FALSE)  
summary(m.full) # at-issueness subsumes all other effects except for arousal, but this needs to be investigated much more to test causal hypotheses about how these factors interact

m.valar <- lmer(response_num ~ carousal_mean*crelativeValence + (1 | Word) + (1 | content) + (1|workerid), data = data, REML = FALSE)  
summary(m.valar) # 

```

## BERT
```{r, include = FALSE, warning = FALSE, echo=FALSE}
m.bert.CC <- lmer(response_num ~ carousal_mean*crelativeValence + catissue + cmeanBertSurprisalCC + (1 | Word) + (1 | content) + (1|workerid), data = data, REML = FALSE)
summary(m.bert.CC) 

m.bert.NoCC <- lmer(response_num ~ carousal_mean*crelativeValence + catissue + cmeanBertSurprisalNoCC + (1 | Word) + (1 | content) + (1|workerid), data = data, REML = FALSE)
summary(m.bert.NoCC) 
```



### Main analysis including Anand et al Categories

```{r bring in the Anand et al categories}
# data$cFativity = data$factivity - mean(data$factivity)
data$cAgentivity = data$agentivity - mean(data$agentivity)
names(data)
```

```{r mixed effects linear model, include = FALSE, warning = FALSE, echo=FALSE}
m.full.AGH <- lmer(response_num ~ carousal_mean + cSurprisal + crelativeValence + catissue + cAgentivity + broad_class + (1 | Word) + (1 | content) + (1|workerid), data = data, REML = FALSE)  
summary(m.full.AGH) # at-issueness subsumes all other effects except for arousal, but this needs to be investigated much more to test causal hypotheses about how these factors interact

m.full <- lmer(response_num ~ carousal_mean + cSurprisal + crelativeValence + catissue + (1 | Word) + (1 | content) + (1|workerid), data = data, REML = FALSE)
summary(m.full)
```

