---
title: "exp1_analysis"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(ggplot2)
library(jsonlite)
library(tidyr)
library(stringr)
library(dplyr)
library(ggthemes)
library(ggrepel)
library(dichromat)
library(lme4)
library(lmerTest)
library(performance)
library(languageR)
library(gghighlight)
theme_set(theme_bw())
library(brms)
library(rstan)
# library(plot3D)


setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
# color-blind-friendly palette
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7") 
```

```{r import data}
top_words_selected <- read.csv("../explore/top_words_selected.csv")
top_words_selected <- select(top_words_selected, "Word", "voice", "valence_group", "valence_bin", "valence_bin_group", "arousal_bin", "valence_arousal_group", "carousal_mean", "crelativeValence", "valence_mean","mean_residuals_abs","mean_acceptability_norm")
data_raw <- read.csv("../../Submiterator/exp1/exp1-trials.csv")
data_participants_raw <- read.csv("../../Submiterator/exp1/exp1-subject_information.csv")
data_at_issue <- read.csv("../../data/atissue_data-bypred.csv")
frequencies = read_csv("../../data/frequencies.csv")
anand = read_csv("../../data/anand-etal-typology_filledout.csv")
names(anand)[names(anand) == "predicate"] <- "Word"
names(anand)[names(anand) == "broad class"] <- "broad_class"
bert_score = read_csv("../../data/bert_scores.csv")
```

# Native speakers
## We exclude any speakers who did not report English as a native language in the post-survey demographic questions.
```{r filter native English speakers, warning = FALSE}
# get participants whose native language is English
data_participants <- data_participants_raw %>%
  mutate(language = tolower(language))

#filter out participants who did not rep ort English as native language
data_participants = filter(data_participants, grepl("eng", language))

worker_id <- data_participants$workerid

data_select <- data_raw %>%
  filter(utterance != "bot_check") %>%
  filter(workerid %in% worker_id)

data_select <- data_select %>%
  mutate(response_num = as.numeric(response))
  #mutate(response_num = as.numeric(levels(response))[response])
```
# Control trials
## We exclude any speakers whose mean performance on control trials is 3 SD above the group mean.
```{r filter out control trials}
data_control <- data_select %>%
  filter(exp == "control") 

ggplot(data_control, aes(x=response_num)) +
  geom_histogram() +
  facet_wrap(~utterance)

worker_control = ggplot(data_control, aes(x=response_num)) +
  geom_histogram() +
  facet_wrap(~workerid)
# ggsave(worker_control, file="graphs/worker_control_items.pdf",width=15,height=20)

control_mean = mean(data_control[, "response_num"])
control_median = median(data_control[, "response_num"])
control_sd = sd(data_control[, "response_num"])
max_control = control_mean + 3 * control_sd

# omit mean response 3 SD above the group mean (per factive paper)
id_omit <- data_control %>%
  group_by(workerid) %>%
  summarise(response_mean_control = mean(response_num)) %>%
  filter(max_control < response_mean_control)
id_omit <- id_omit[, "workerid"]

id_select <- data_control %>%
  group_by(workerid) %>%
  summarise(response_mean_control = mean(response_num)) %>%
  filter(response_mean_control <= max_control)

id_select <- pull(id_select[c("workerid")])
```

```{r filter out stimuli trials}
data_stim <- data_select %>% 
  filter(workerid %in% id_select) %>%
  filter(exp == "stim") %>%
  mutate(Word = ifelse(Word == "facinate", "fascinate", as.character(Word))) ## coded "fascinate" wrongly :(

data <- merge(data_stim, top_words_selected, by = "Word")

data$valence_arousal_group = factor(data$valence_arousal_group, levels=c("high negative valence, low arousal", "high negative valence, medium arousal", "high negative valence, high arousal", "low valence, low arousal", "low valence, medium arousal", "low valence, high arousal", "high positive valence, low arousal", "high positive valence, medium arousal", "high positive valence, high arousal")) 

predlevels = unique(data$Word) %>% 
  sort()
# JD: hm, "confess" and "remark" are in frequencies but not in data, wonder why?
frequencies = frequencies %>% 
  filter(! Word %in% c("confess","remark")) %>% 
  arrange(Word)
frequencies$Word = predlevels # risky new assignment if predicates not ordered correctly
data = data %>% 
  left_join(data_at_issue,by=c("Word")) %>% 
  left_join(frequencies, by=c("Word"))

# Add in the Anand et al (2019) coding
data = left_join(data,anand, by=c("Word"))

bert_score$BertSurprisal = -log(bert_score$BertScore)
avg_bert = bert_score %>%
  group_by(Verb) %>%
  summarize(meanBertSurprisal = mean(BertSurprisal))

names(avg_bert)[names(avg_bert) == "Verb"] <- "Word"
data = left_join(data,avg_bert, by=c("Word"))

View(data)
```

```{r average by word}
data_by_word <- data %>%
  group_by(Word, voice, valence_group, valence_bin, valence_bin_group, arousal_bin, valence_arousal_group, carousal_mean, crelativeValence,mean_atissue,Surprisal,BertSurprisal,broad_class) %>%
  summarise(response_mean = mean(response_num))
View(data)
```

# Plots

## Unnormalized projection rating against relative valence

```{r}
ggplot(data = data_by_word, aes(x = crelativeValence, y = response_mean, label = Word)) +
  geom_point(width = .3,height = .025) +
  geom_smooth(method = 'lm') + 
  geom_text_repel() +
  labs (x = "Relative predicate valence", y = "Mean projection rating")
ggsave("graphs/means_by_valence.pdf",width=4,height=3)

```

## Unnormalized rating agaist arousal mean
```{r}
ggplot(data = data_by_word, aes(x = carousal_mean, y = response_mean, label = Word)) +
  geom_point(width = .3,height = .025) +
  geom_text_repel() +
  geom_smooth(method = 'lm') +
  labs (x = "Relative predicate arousal", y = "Mean projection rating")
ggsave("graphs/means_by_arousal.pdf",width=4,height=3)
```

## Unnormalized projection rating against relative valence, by valence group

```{r}
ggplot(data = data_by_word, aes(x = crelativeValence, y = response_mean, label = Word, color=valence_group)) +
  geom_point(width = .3,height = .025) +
  geom_smooth(method = 'lm') + 
  geom_text_repel() +
  scale_color_manual(values=cbPalette,name="Valence") +
  labs (x = "Relative predicate valence", y = "Mean projection rating")
ggsave("graphs/means_by_valence.pdf",width=4.5,height=3)

```

## Unnormalized rating agaist arousal mean, by valence group
```{r}
ggplot(data = data_by_word, aes(x = carousal_mean, y = response_mean, label = Word,color=valence_group)) +
  geom_point(width = .3,height = .025) +
  geom_text_repel() +
  geom_smooth(method = 'lm') +  
  scale_color_manual(values=cbPalette,name="Valence") +
  labs (x = "Relative predicate arousal", y = "Mean projection rating")
ggsave("graphs/means_by_arousal.pdf",width=4.5,height=3)
```


## Unnormalized rating agaist at-issueness

```{r}
ggplot(data = data_by_word, aes(x = mean_atissue, y = response_mean, label = Word)) +
  geom_point(width = .3,height = .025) +
  geom_text_repel() +
  geom_smooth(method = 'lm') + 
  labs (x = "Mean at-issueness rating", y = "Mean projection rating", colour = "predicate group")
ggsave("graphs/projection_by_atissue.pdf",width=4.5,height=3.5)

```

## Verb classes
```{r}

ggplot(data = data_by_word, aes(x = crelativeValence, y = response_mean, label = Word)) +
  geom_point(width = .3,height = .025) +
  geom_smooth(method = 'lm') + 
  geom_text_repel() +
  facet_wrap(~broad_class) +
  labs (x = "Relative predicate valence", y = "Mean projection rating")
ggsave("graphs/means_broad_class_by_valence.pdf",width=4,height=3)


# uncenter these values
ggplot(data = data_by_word, aes(x = crelativeValence, fill = broad_class)) +
  geom_density(alpha = .4) +
  labs (x = "Relative predicate valence", y = "Mean projection rating")

ggplot(data = data_by_word, aes(x = carousal_mean, y = response_mean, label = Word)) +
  geom_point(width = .3,height = .025) +
  geom_text_repel() +
  facet_wrap(~broad_class) +
  geom_smooth(method = 'lm') +
  labs (x = "Relative predicate arousal", y = "Mean projection rating")
ggsave("graphs/means_broad_class_by_arousal.pdf",width=4,height=3)

ggplot(data = data_by_word, aes(x = carousal_mean, fill = broad_class)) +
  geom_density(alpha = .4) +
  labs (x = "Relative predicate arousal", y = "Mean projection rating")




ggplot(data_by_word, aes(x = carousal_mean, y = response_mean, fill = broad_class)) +
  labs (x = "Relative predicate arousal", y = "Mean projection rating")


ggplot(data = data_by_word, aes(x = mean, y = carousal_mean, label = Word)) +
  geom_point(width = .3,height = .025) +
  geom_text_repel() +
  facet_wrap(~broad_class) +
  geom_smooth(method = 'lm') +
  labs (x = "Relative predicate arousal", y = "Mean projection rating")
```


```{r}

ggplot(data = data_by_word, aes(x = carousal_mean, y = response_mean, label = Word,color=valence_group)) +
  geom_point(width = .3,height = .025) +
  geom_text_repel() +
  geom_smooth(method = 'lm') +  
  scale_color_manual(values=cbPalette,name="Valence") +
  labs (x = "Relative predicate arousal", y = "Mean projection rating")
ggsave("graphs/means_by_arousal.pdf",width=4.5,height=3)
```


```{r, include = FALSE}
scatter3D(data_by_word$crelativeValence, data_by_word$carousal_mean, data_by_word$response_mean, theta = 30, phi = 10,
          xlab = "relative valence",
          ylab ="arousal mean", 
          zlab = "response mean")
```


```{r extra plots, include = FALSE}
# some extra plots
ggplot(data = data_by_word, aes(x = crelativeValence, y = response_mean, colour = valence_group, label = Word)) +
  geom_point(width = .3,height = .025) +
  geom_label() +
  geom_smooth(method = 'lm') +
  scale_color_manual(values=cbPalette)

ggplot(data = data_by_word, aes(x = carousalMean, y = response_mean, colour = valence_group, label = Word)) +
  geom_point(width = .3,height = .025) +
  geom_label() +
  geom_smooth(method = 'lm') +
  scale_color_manual(values=cbPalette)

ggplot(data = data_by_word, aes(x = crelativeValence, y = response_mean, colour = valence_group, label = Word)) +
  geom_point(width = .3,height = .025) +
  facet_grid(~voice) +
  geom_label() +
  geom_smooth(method = 'lm')

ggplot(data = data_by_word, aes(x = crelativeValence, y = response_mean, colour = valence_group, label = Word)) +
  geom_point(width = .3,height = .025) +
  facet_grid(~valence_arousal_group) +
  geom_label() +
  geom_smooth(method = 'lm')

#maybe color with gradient? 
#maybe color = arousal, shape = valence? 
ggplot(data = data_by_word, aes(x = carousal_mean, y = response_mean, colour = arousal_bin, label = Word)) +
  geom_point(width = .3,height = .025) +
  geom_label() +
  geom_smooth(method = 'lm')

ggplot(data = data_by_word, aes(x = carousal_mean, y = response_mean, colour = arousal_bin, label = Word)) +
  geom_point(width = .3,height = .025) +
  facet_grid(~voice) +
  geom_label() +
  geom_smooth(method = 'lm')

ggplot(data = data_by_word, aes(x = carousal_mean, y = response_mean, colour = arousal_bin, label = Word)) +
  geom_point(width = .3,height = .025) +
  facet_grid(~valence_arousal_group) +
  geom_label() +
  geom_smooth(method = 'lm')
```


