---
title: "Categorical Timed Expt Analysis"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(ggplot2)
library(jsonlite)
library(tidyr)
library(stringr)
library(dplyr)
library(ggthemes)
library(ggrepel)
library(dichromat)
library(lme4)
library(lmerTest)
library(performance)
library(languageR)
library(gghighlight)
theme_set(theme_bw())
library(brms)
library(rstan)
library(plot3D)
```

```{r import data}
top_words_selected <- read.csv("../explore/top_words_selected.csv")
top_words_selected <- select(top_words_selected, "Word", "voice", "valence_group", "valence_bin", "valence_bin_group", "arousal_bin", "valence_arousal_group", "carousal_mean", "crelativeValence", "valence_mean", "mean_residuals_abs","mean_acceptability_norm")
data_raw <- read.csv("../../data/2_categoricaltimed/study-trials.csv")
data_participants_raw <- read.csv("../../data/2_categoricaltimed/study-subject_information.csv")
#data_at_issue <- read.csv("../../data/atissue_data-bypred.csv")

data_at_issue <- read.csv("../../data/atissue_data-bypred.csv")
frequencies = read_csv("../../data/frequencies.csv")
anand = read_csv("../../data/anand-etal-typology_filledout.csv")
names(anand)[names(anand) == "predicate"] <- "Word"
names(anand)[names(anand) == "broad class"] <- "broad_class"
bert_score = read_csv("../../data/bert_scores.csv")
```

# Native speakers
## We exclude any speakers who did not report English as a native language in the post-survey demographic questions.
```{r filter native English speakers, warning = FALSE}
# get participants whose native language is English
data_participants <- data_participants_raw %>%
  mutate(language = tolower(language))

#filter out participants who did not report English as native language
data_participants = filter(data_participants, grepl("eng", language))

worker_id <- data_participants$workerid

data_select <- data_raw %>%
  filter(utterance != "bot_check") %>%
  filter(workerid %in% worker_id)

data_select <- data_select %>%
  mutate(response_num = as.numeric(response))
  #mutate(response_num = as.numeric(levels(response))[response])
```
# Control trials
## We exclude any speakers whose response to more than 2 out of 6 controls is "Yes".
```{r filter out control trials}
data_control <- data_select %>%
  filter(exp == "control") 

id_select <- data_control %>%
  group_by(workerid) %>%
  summarise(yes_response = sum(response == "yes")) %>%
  filter(yes_response <= 2)
id_select <- id_select[,"workerid"]

#exclude 5 ppts

id_select <- pull(id_select[c("workerid")])
```

```{r filter out stimuli trials}
data_stim <- data_select %>% 
  filter(workerid %in% id_select) %>%
  filter(exp == "stim") %>%
  filter(rt != "Late") %>%
  mutate(Word = ifelse(Word == "facinate", "fascinate", as.character(Word))) ## coded "fascinate" wrongly :( 

data <- merge(data_stim, top_words_selected, by = "Word")

data$valence_arousal_group = factor(data$valence_arousal_group, levels=c("high negative valence, low arousal", "high negative valence, medium arousal", "high negative valence, high arousal", "low valence, low arousal", "low valence, medium arousal", "low valence, high arousal", "high positive valence, low arousal", "high positive valence, medium arousal", "high positive valence, high arousal")) 
```

```{r average by word, coded 1 for yes, 0 for no}
data_by_word <- data %>%
  group_by(Word, voice, valence_group, valence_bin, valence_bin_group, arousal_bin, valence_arousal_group, carousal_mean, crelativeValence) %>%
  mutate(response_num = ifelse(response == "yes", 1, 0)) %>%
  summarise(response_mean = mean(response_num))
```

# Plots
## Unnormalized rating agaist relative valence

```{r}
ggplot(data = data_by_word, aes(x = crelativeValence, y = response_mean, label = Word)) +
  geom_point(width = .3,height = .025, aes(colour = valence_arousal_group)) +
  geom_text_repel(aes(colour = valence_arousal_group)) +
  geom_smooth(method = 'lm') + 
  labs (x = "relative valence", y = "mean projectivity rating", colour = "predicate group")
```

## Unnormalized rating agaist arousal mean
```{r}
ggplot(data = data_by_word, aes(x = carousal_mean, y = response_mean, label = Word)) +
  geom_point(width = .3,height = .025, aes(colour = valence_arousal_group)) +
  geom_text_repel(aes(colour = valence_arousal_group)) +
  geom_smooth(method = 'lm') +   
  labs (x = "mean arousal", y = "mean projectivity rating", colour = "predicate group")
```


```{r extra plots, include = FALSE}
# some extra plots
ggplot(data = data_by_word, aes(x = crelativeValence, y = response_mean, colour = valence_group, label = Word)) +
  geom_point(width = .3,height = .025) +
  geom_label() +
  geom_smooth(method = 'lm')

ggplot(data = data_by_word, aes(x = crelativeValence, y = response_mean, colour = valence_group, label = Word)) +
  geom_point(width = .3,height = .025) +
  facet_grid(~voice) +
  geom_label() +
  geom_smooth(method = 'lm')

ggplot(data = data_by_word, aes(x = crelativeValence, y = response_mean, colour = valence_group, label = Word)) +
  geom_point(width = .3,height = .025) +
  facet_grid(~valence_arousal_group) +
  geom_label() +
  geom_smooth(method = 'lm')

#maybe color with gradient? 
#maybe color = arousal, shape = valence? 
ggplot(data = data_by_word, aes(x = carousal_mean, y = response_mean, colour = arousal_bin, label = Word)) +
  geom_point(width = .3,height = .025) +
  geom_label() +
  geom_smooth(method = 'lm')

ggplot(data = data_by_word, aes(x = carousal_mean, y = response_mean, colour = arousal_bin, label = Word)) +
  geom_point(width = .3,height = .025) +
  facet_grid(~voice) +
  geom_label() +
  geom_smooth(method = 'lm')

ggplot(data = data_by_word, aes(x = carousal_mean, y = response_mean, colour = arousal_bin, label = Word)) +
  geom_point(width = .3,height = .025) +
  facet_grid(~valence_arousal_group) +
  geom_label() +
  geom_smooth(method = 'lm')
```



# Mixed effects linear model and ANOVA
```{r, include = FALSE, warning = FALSE, echo=FALSE}
# data_model <- data
#
m = glm(response ~ carousal_mean * crelativeValence + (1 + (carousal_mean * crelativeValence) | workerid) + (1 | Word) + (1 + (carousal_mean * crelativeValence) | content), data = data_model, family = "binomial")
# summary(m) # main effects of arousal and relative valence!
# plot(fitted(m), residuals(m))

#pdf(file="modelcheck_linear.pdf",height=8,width=9)
#check_model(m)
#dev.off()
```

```{r anova, include = FALSE, warning = FALSE, echo=FALSE}
# projectivity.full = lmer(response_num ~ carousal_mean * crelativeValence + (1 + (carousal_mean * crelativeValence) | workerid) + (1 | Word) + (1 + (carousal_mean * crelativeValence) | content), data = data_model, REML = FALSE)
# projectivity.reduced = lmer(response_num ~ 1 + (1 + (carousal_mean * crelativeValence) | workerid) + (1 | Word) + (1 + (carousal_mean * crelativeValence) | content), data = data_model, REML = FALSE)
# anova(projectivity.reduced, projectivity.full)
# 
# 
# #interaction removed
# projectivity.int = lmer(response_num ~ carousal_mean + crelativeValence + (1 + (carousal_mean * crelativeValence) | workerid) + (1 | Word) + (1 + (carousal_mean * crelativeValence) | content), data = data_model, REML = FALSE)
# anova(projectivity.int, projectivity.full)
# 
# #only arousal
# projectivity.arousal = lmer(response_num ~ carousal_mean + (1 + (carousal_mean * crelativeValence) | workerid) + (1 | Word) + (1 + (carousal_mean * crelativeValence) | content), data = data_model, REML = FALSE)
# anova(projectivity.int, projectivity.arousal)
# 
# #only valence
# projectivity.valence = lmer(response_num ~ crelativeValence + (1 + (carousal_mean * crelativeValence) | workerid) + (1 | Word) + (1 + (carousal_mean * crelativeValence) | content), data = data_model, REML = FALSE)
# anova(projectivity.int, projectivity.valence)

```

```{r mixed bayes, include = FALSE, echo = FALSE}

# # m.bayes = brm(response_num ~ carousal_mean * crelativeValence + (1 + (carousal_mean * crelativeValence) | workerid) + (1 | Word) + (1 + (carousal_mean * crelativeValence) | content), 
# #               data = data_model,
# #               chains = 4,
# #               cores = 4)
# # saveRDS(m.bayes, "bayes.rds")
# m.bayes <- readRDS("bayes.rds")
# summary(m.bayes) # 0 is in 95% CI for interaction, but not for valence / arousal separately
```


```{r bring in the at-issueness data to test effect on projection, include = FALSE}

# View(data_at_issue)
# # subset to just words and mean at-issue rating
# # rename mean_response column from at-issue data
# names(data_at_issue)[names(data_at_issue) == "cresponse_mean_pred"] <- "catissueness"
# 
# data_proj <- data %>%
#   group_by(Word,content)%>%
#   summarize(mean_projection = mean(response_num))
# nrow(data_proj) #1908
# 
# # merge together
# total = left_join(data_proj,data_at_issue)
# 
# # Add crelativeValence and carousal_mean
# total = left_join(total,data_by_word[,c("Word","carousal_mean","crelativeValence")])
# View(total)
```

# lmer(projection_response ~ cvalence*carousal*catissueness...)
```{r, include = FALSE, warning = FALSE, echo=FALSE}

# m.proj.atissue.full <- lmer(mean_projection ~ carousal_mean*crelativeValence*catissueness + (1 | Word) + (1 + (carousal_mean*crelativeValence) | content), data = total, REML = FALSE)  
# 
# summary(m.proj.atissue.full) # 
# plot(fitted(m.proj.atissue.full), residuals(m.proj.atissue.full))


```

```{r}
predlevels = unique(data$Word) %>% 
  sort()
# JD: hm, "confess" and "remark" are in frequencies but not in data, wonder why?
frequencies = frequencies %>% 
  filter(! Word %in% c("confess","remark")) %>% 
  arrange(Word)
frequencies$Word = predlevels # risky new assignment if predicates not ordered correctly
data = data %>% 
  left_join(data_at_issue,by=c("Word")) %>% 
  left_join(frequencies, by=c("Word"))

# Add in the Anand et al (2019) coding
data = left_join(data,anand, by=c("Word"))

bert_score$BertSurprisal = -log(bert_score$BertScore)
avg_bert = bert_score %>%
  group_by(Verb) %>%
  summarize(meanBertSurprisal = mean(BertSurprisal))

names(avg_bert)[names(avg_bert) == "Verb"] <- "Word"
data = left_join(data,avg_bert, by=c("Word"))
```



Main analysis reported by JD in workshop talk

```{r bring in the at-issueness data to test effect on projection}
names(data)
data$workerid = as.factor(as.character(data$workerid))
data$catissue = data$mean_atissue - mean(data$mean_atissue)
# Surprisal from google n-gram
data$cSurprisal = data$Surprisal - mean(data$Surprisal)
View(data$BertSurprisal)

# surprisal from BERT INCLUDING CC
data$cBert = data$meanBertSurprisal - mean(data$mean_bert_score)
data$cBertSurprisal = data$BertSurprisal - mean(data$BertSurprisal)
```

```{r mixed effects linear model, include = FALSE, warning = FALSE, echo=FALSE}
m.full <- lmer(response_num ~ carousal_mean*crelativeValence + catissue + cSurprisal + (1 | Word) + (1 | content) + (1|workerid), data = data, REML = FALSE)  
summary(m.full) # at-issueness subsumes all other effects except for arousal, but this needs to be investigated much more to test causal hypotheses about how these factors interact

m.valar <- lmer(response_num ~ carousal_mean*crelativeValence + (1 | Word) + (1 | content) + (1|workerid), data = data, REML = FALSE)  
summary(m.valar) # 

```
```{r}

```

### Main analysis using bert scores rather than surprisal from google books
```{r Using surprisal estimates from BERT in a masked language modeling task, include = FALSE, warning = FALSE, echo=FALSE}
m.bert <- lmer(response_num ~ carousal_mean*crelativeValence + catissue + cBertSurprisal + (1 | Word) + (1 | content) + (1|workerid), data = data, REML = FALSE)
summary(m.bert) 

# exploratory
m.bert <- lmer(response_num ~ carousal_mean*crelativeValence*broad_class + (1 | Word) + (1 | content) + (1|workerid), data = data, REML = FALSE)
summary(m.bert) 

m.bert <- lmer(response_num ~ carousal_mean*crelativeValence + catissue + broad_class + cBertSurprisal + (1 | Word) + (1 | content) + (1|workerid), data = data, REML = FALSE)
summary(m.bert) 
```

### Main analysis including Anand et al Categories

```{r bring in the Anand et al categories}
# data$cFativity = data$factivity - mean(data$factivity)
data$cAgentivity = data$agentivity - mean(data$agentivity)
names(data)
```

```{r mixed effects linear model, include = FALSE, warning = FALSE, echo=FALSE}
m.full.AGH <- lmer(response_num ~ carousal_mean + cSurprisal + crelativeValence + catissue + cAgentivity + broad_class + (1 | Word) + (1 | content) + (1|workerid), data = data, REML = FALSE)  
summary(m.full.AGH) # at-issueness subsumes all other effects except for arousal, but this needs to be investigated much more to test causal hypotheses about how these factors interact

m.full <- lmer(response_num ~ carousal_mean + cSurprisal + crelativeValence + catissue + (1 | Word) + (1 | content) + (1|workerid), data = data, REML = FALSE)
summary(m.full)
```



